---
title: "Elderly Care Analysis"
author: "Michael"
date: "November 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Healthcare & the Elderly 
Our question involves disparity in the care for the elderly based on income and racial distrbutrions in California. We will be using spatial analysis to determine if there is a spatial pattern based on the dispersion (or lackthereof) of care facilities based on their bed size/capacity size. 

We will be testing the hypothesis that facility capacity is correlated with income and race. There is also reason to believe that location matters: areas with neighboring low-income, POC populations will correspond with less bed sizes. 

To test this, we will need to test if there is a spatial pattern, and then carry on to run some spatial regression models. 

##Data Prep
We will start with two basic datasets provided by Jen on the elderly/adult care facilities in CA. 

```{r, include=FALSE, echo=FALSE}
library(tidyverse)
library(ggmap)
library(tidycensus)

register_google(key = "AIzaSyDFKaPxRWg92bYqtj2xRjQcVoq-LAgI8dU", account_type = "premium", day_limit = 100000)
ggmap_credentials()

register_google(key = "AIzaSyDFKaPxRWg92bYqtj2xRjQcVoq-LAgI8dU")
```

```{r, echo=FALSE, eval=FALSE}
adult.res <- read_csv("AdultResidentialFacilities10292017.csv", col_names = TRUE)
elderly.res <- read_csv("ResidentialElderCareFacility10292017.csv", col_names = TRUE)

```

Things are looking a little messy, so let's clean up the dataset so we only keep the variables of interest. 
```{r, eval=F}
adult.prep <- adult.res %>%
  select(Type, Name, Administrator, Address, City, State, Zip, `County Name`, Capacity, `License First Date`, Status) %>% 
  filter(Type == "ADULT RESIDENTIAL" & Status == "LICENSED") 

elderly.prep <- elderly.res %>%
  select(Type, Name, Administrator, Address, City, State, Zip, `County Name`, Capacity, `License First Date`, Status) %>%
  filter(Type == "RESIDENTIAL CARE ELDERLY" & Status == "LICENSED")

#Let's join the two 
care.facilities <- adult.prep %>%
  full_join(elderly.prep)

write_csv(care.facilities, "facilities.prepped.csv")
```


Now we can just load the prepped csv, and begin working on it. 
* First, we want to get combine the address, city, state, zip into one address 
* Then we can send that to google maps api to get the geo coordinates for mapping 

```{r, eval=FALSE}
care.facilities <- read.csv("facilities.prepped.csv")
care.facilities$Location <- paste0(care.facilities$Address, ", ", care.facilities$City, ", ", care.facilities$State, care.facilities$Zip)


geo <- geocode(location = care.facilities$Location, output="latlon", source="google")

care.facilities$Lat <- geo$lat
care.facilities$Long <- geo$lon
```

Adding in census data. 
```{r, eval=FALSE}
census_api_key(Sys.getenv("CENSUS_API_KEY"))

care.facilities <- read_csv("care.fac.geo.csv")

#get variables 
census_var <- load_variables(2015, "acs5", cache = TRUE)

test <- get_acs(geography = "tract", variables = "B06012_001E", state="CA", geometry = TRUE)


#getting data based county 
ca.income.county <- get_acs(geography = "county", variables = "B06011_001E", state = "CA", geometry = TRUE)
```

Mapping
```{r}
library(leaflet)
library(stringr)
library(sf)
library(viridis)
library(viridisLite) 

care.facilities <- read_csv("care.fac.geo.csv")

#getting data on median income, census tract
ca.house.income <- get_acs(geography = "tract", variables = "B19013_001", state = "CA", geometry = TRUE)

ca.home.value <- get_acs(geography = "tract", variables = "B25077_001", state = "CA", geometry = TRUE)

#setting color 
pal <- colorNumeric(palette = "viridis", 
                    domain = ca.house.income$estimate)

getColor <- function(care.facilities) {
  sapply(care.facilities$Capacity, function(Capacity) {
  if(Capacity <= 12) {
    "red"
  } else if(Capacity >= 13 & Capacity <= 24) {
    "orange"
  } else if(Capacity >= 25 & Capacity <=100) {
    "yellow"
  } else {
    "green"
  } } )
}


ca.map <- ca.house.income %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
      addTiles() %>%
      addPolygons(stroke= FALSE,
                smoothFactor = 0,
                fillOpacity = 0.7,
                color = ~ pal(estimate)) %>%
      addCircles(data = care.facilities, 
                 lat = ~Lat,
                 lng =  ~Long,
                 label = ~as.character(Name),
                 weight = 2, 
                 radius = 20,
                 stroke = TRUE,
                 fillOpacity = 0.8,
                 color = getColor(care.facilities)
                 ) %>%
      addLegend("bottomright", 
              pal = pal, 
              values = ~ estimate,
              title = "Household Income",
              labFormat = labelFormat(prefix = "$"),
              opacity = 1) %>%
      addLegend("bottomleft", 
                colors= c("Red", "Orange", "Yellow", "Green"),
                labels = c("4-12 Beds", "13-24 Beds", "25-100 Beds", "101+ Beds"),
                title = "Capacity Size")


```
Okay so for the most part, I figured it out, no idea why there are some corrdinates in like Michigan. Time to aggregate data for regression analysis. 

What I'm thinking is we can try to map all the points and see what it looks like first. Using circle radius perhaps? Checklist of things: 

* Map of all the locations on top of income 
* graph of amount of buildings over time - years 
* If Time: Spatial regression 

```{r}
#save the map 
library(htmlwidgets)

saveWidget(ca.map, "ca.map.html", selfcontained = FALSE)

ca.map
```

Stuff over time: 
```{r}
library(lubridate)

care.facilities$date <- mdy(care.facilities$License.First.Date)

care.facilities$year <- year(care.facilities$date)

levels <- c(-Inf, 12, 25, 100, Inf)
labels <- c("small", "medium", "large", "x-large")

care.facilities$size <- cut(care.facilities$Capacity, levels, labels)

#time to re-organize data for graphing 
growth <- care.facilities %>%
  select(size, year) %>%
  group_by(year) %>%
  count(size)

growth.plot <- ggplot(growth, aes(x=year, y=n, color=size)) + 
  geom_line()

growth.plot <- growth.plot + ggtitle("Growth of Care Facilities, CA") +
  xlab("Year") +
  ylab("Facilities Opened")

growth.plot
```
Okay, so now we need to prepare data for spatial regression. We can accomplishment this by aggregating the data in a way that satisfies our questions: Is facility size influenced by income? Or maybe race? 

To do this, let's create a dummy variable that gives us a "1" if the facility is small, and a "0" if it is not. The amount of small facilities in a census blocked will be our dependent variable in our regression models. 

```{r}
care.facilities <- care.facilities %>%
  mutate(dummy = ifelse(size == "small", 1, 0))

#create a new csv with just the variables we need 

care.regression <- care.facilities %>%
  select(Name, City, Lat, Long, dummy)

write.csv(care.regression, "care.regression.csv")

#playing around with the data 
care.reg.agg <- care.regression %>%
  select(City, dummy) %>%
  group_by(City) %>%
  summarise(small = sum(dummy == 1))

#It might be possible to group by city and do a proportion? 
care.city <- care.facilities %>%
  select(City, size) %>%
  group_by(City) %>%
  summarise(small = sum(size == "small"),
            other = sum(size != "small"),
            ratio = small/other)

care.city.measure <- care.facilities %>%
  select(City, size) %>%
  group_by(City) %>%
  count(size)

ca.house.income <- get_acs(geography = "tract", variables = "B19013_001", state = "CA", geometry = TRUE)

```
