---
title: "Elderly Care Analysis"
author: "Michael"
date: "November 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Healthcare & the Elderly 
Our question involves disparity in the care for the elderly based on income and racial distrbutrions in California. We will be using spatial analysis to determine if there is a spatial pattern based on the dispersion (or lackthereof) of care facilities based on their bed size/capacity size. 

We will be testing the hypothesis that facility capacity is correlated with income and race. There is also reason to believe that location matters: areas with neighboring low-income, POC populations will correspond with less bed sizes. 

To test this, we will need to test if there is a spatial pattern, and then carry on to run some spatial regression models. 

##Data Prep
We will start with two basic datasets provided by Jen on the elderly/adult care facilities in CA. 

```{r, include=FALSE}
library(tidyverse)
```
```{r, echo=FALSE}
adult.res <- read_csv("AdultResidentialFacilities10292017.csv")
elderly.res <- read_csv("ResidentialElderCareFacility10292017.csv")

```
Things are looking a little messy, so let's clean up the dataset so we only keep the variables of interest. 
```{r}
adult.prep <- adult.res %>%
  select(`Facility Type`, `Facility Name`, `Facility Administrator`, `Facility Address`, `Facility City`, `Facility State`, `Facility Zip`, Faci)
